### 2016-07-01 image_setup.tex
Disclaimer about proper copied characters rendering in output PDF

\label{sec:image_setup}

\subsection{Installing Raspbian}

To get started, we first need to install an \ac{OS} on the
\ac{Raspi}s. We will use the popular Debian-based Raspbian
Linux~\cite{raspbian}

We will be using a \ac{PC} running a Debian-based Linux distribution
to setup a customized Raspbian Linux image that can be written to
a memory card and work in any \ac{Raspi} version 1 or 2.

Raspbian is made available in two bundles: Raspbian and Raspbian lite.
The difference between the two is that Raspbian contains a pre-installed desktop
environment, and Raspbian lite is by default only accessible through a 
command shell. Both Raspbian and Raspbian lite will both by default be remotely
accessible using \ac{SSH}. \ac{SSH} provides the means to remotely
connect to another machine through a command shell.
%In fact it is the successor
%of \ac{Telnet} that has been designed with network security in mind.

The \ac{Raspi}s in our testbed are not connected to monitors, so we
have decided to go with Raspbian lite.
%Our testbed does not have monitors so we will install Raspbian lite.
A desktop environment can be installed using the package manager
(apt-get) in case a desktop environment will be desired at a later
time.

Below is the procedure we followed to setup the \ac{Raspi}s in our testbed. We
have performed a simple setup where each \ac{Raspi} is equipped with a
memory card that contains a slightly customized Raspbian lite image. The image
of the \ac{OS} in all \ac{Raspi}s will be identical.

The overall procedure to customize the official Raspbian lite image
are:
\begin{enumerate}
    \item Download Raspbian lite
    \item Alter Raspbian lite. e.g. browse, modify, add, and delete files
        in the official Raspbian lite image
    \item Change root, i.e. change root filesystem into the Raspbian lite
        image to update and install software packages
    \item Write image to memory cards
\end{enumerate}

Hashmark (\#) and dollar (\$) will be used throughout the code blocks in
the paper to indicate if an operation needs to be run with root permissions or
user permissions.

\subsection{Download Raspbian lite}

The latest Raspbian lite bundle can be downloaded from Raspbian's official
webpage~\cite{raspbian}.
%To download the latest Raspbian lite, we go to the url
%\url{http://downloads.raspberrypi.org/raspbian_lite/images/}.
At the time of this writing, the latest available bundle was
\texttt{2016-05-27-raspbian-jessie-lite.zip}.

To ensure that the content of the bundle does not change, this procedure
will be based on that particular version of Raspbian lite which we have
made available at~\cite{tunescode_webpage}. All other files used in this paper
will also be available there unless specified otherwise.

%We want to proceed downloading the bundle in a shell instead of the browser.
%To do this, first open a command shell on a linux machine. This guide is
%assuming a Debian-based distribution, but there should only be minor differences
%if you have another package manager. i.e. the package managers and name of software packages.

To get started, open a Linux shell (terminal) and declare the variables below.
The \texttt{URL} and \texttt{IMAGE} specifies where the Linux bundle will be
located and \texttt{WORKDIR} specifies a working directory where we will 
store Raspbian and customize it.
Do not type in \$ and \# in the shell. Those indicates if a command can be
executed as a regular user or if root permissions are required.
Running a command as root can on most systems be done by putting
\texttt{sudo} in front of the command. This is illustrated in the
following code block 
with the command \texttt{whoami} that prints the username. Lines within a code
block without leading \$ or \# is terminal output or content of a file.

@@

%A final thing to notice is the way \$ and \# are used in the code blocks. \$
%means that the command can be run with normal user permissions where \# indicates that the
%command needs root permissions.
%Running a command with root permissions can on most systems be done by putting
%the command "sudo" in front of each line. This is illustrated in the code below
%with the command "whoami" that prints the username of the caller. Lines without
%leading \$ or \# is the terminal output.

% Root example
\begin{lstlisting}[]
$ whoami
<USERNAME>
$ sudo whoami
root
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Next, make the working directory and used \texttt{cd} to change into the working
directory. Then \texttt{wget} can be used to download Raspbian lite before
unpacking the \texttt{zip} file:

@@

\subsection{The image}

>>>>

\label{sec:image_setup}
In this Section, we review the steps to create a common \ac{OS} image for all the
\ac{Raspi}s. The image setup is composed of three major steps: Select and download the
\ac{OS} image file, alter the image structure and configure the \ac{OS} files.
We proceed to detail all these steps providing brief discussions to our setup
choices when required. To perform these steps, we indicate with command-line
blocks the required sequential commands to be typed by the testbed
administrator in its \ac{PC} to obtain the desired setting. In all the
following sections, the number sign (\#) and dollar sign (\$) will be used in
the command blocks in the paper to indicate if an operation needs
to be run with root permissions or common user permissions, respectively. The
signs will be posted as the terminal prompt in the command blocks.

\subsection{OS Selection and Download}

To get started, we first need to install an \ac{OS} that works properly
on all the \ac{Raspi} models. We will download and setup the image in the testbed
administrator \ac{PC} using a Debian-based distribution. We use the popular
Debian-based Raspbian Linux~\cite{raspbian} given that is the recommended
and default \ac{OS} for the \ac{Raspi}.
Raspbian is made available in two bundles: Raspbian and Raspbian lite.
The difference between the two, is that Raspbian contains a pre-installed desktop
environment for user interaction and Raspbian lite by default only permits to
interact through a command shell. Given that the \ac{Raspi}s in our testbed are
not connected to monitors, we decide to work with Raspbian lite. Still,
if later desired, a desktop environment can be installed using the package manager
as it will be shown in the configuration section.

% The overall procedure to customize the official Raspbian lite image
% are:
% \begin{enumerate}
%     \item Download Raspbian lite
%     \item Alter Raspbian lite. e.g. browse, modify, add, and delete files
%         in the official Raspbian lite image
%     \item Change root, i.e. change root filesystem into the Raspbian lite
%         image to update and install software packages
%     \item Write image to memory cards
% \end{enumerate}

%\subsection{Download Raspbian lite}
%To download the latest Raspbian lite, we go to the url
%\url{http://downloads.raspberrypi.org/raspbian_lite/images/}.

The latest Raspbian lite bundle can be downloaded from the Raspbian
official webpage~\cite{raspbian}. At the time of this writing, the latest
available bundle was \texttt{2016-05-27-raspbian-jessie-lite.zip}.
To ensure that the content of the bundle does not change, this procedure
is based on that particular version of Raspbian lite which we have
made available at~\cite{tunescode_webpage}. All other files used in this paper
are available there unless specified otherwise. To get started, open a
Linux shell (terminal) and declare the environment variables.
To copy the commands from this procedure to the testbed administrator terminal,
we strongly recommend to read this document with Adobe Acrobat Reader in Linux.
Otherwise, copied characters might not render properly.

@@

In the code block, the \texttt{URL} and \texttt{IMAGE} specify where
the Linux bundle is located and \texttt{WORKDIR} specifies a working
directory where the Raspbian bundle will be downloaded and customized.
Notice that even though we use the \$ and \# signs in the shell, in general
these signs will be particular to the testbed administrator \ac{OS} shell.
Next, create the working directory and go to it with the \texttt{cd} command.
To download the image, utilize \texttt{wget} before
unpacking the \texttt{zip} file as follows:

% Running a command as root can on most systems be done by putting
% \texttt{sudo} in front of the command. This is illustrated in the
% following code block 
% with the command \texttt{whoami} that prints the username. Lines within a code
% block without leading \$ or \# is terminal output or content of a file.

% Root example
% \begin{lstlisting}[]
% $ whoami
% <USERNAME>
% $ sudo whoami
% root
% \end{lstlisting}
% \FloatBarrier
% \vspace{-5mm}

@@

\subsection{Image Customization}
